---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# dabestr

dabestr is a package for **D**ata **A**nalysis using **B**ootstrap-Coupled **EST**imation in R.

## About

[Estimation statistics](https://en.wikipedia.org/wiki/Estimation_statistics "Estimation Stats on Wikipedia") is a [simple framework](https://thenewstatistics.com/itns/ "Introduction to the New Statistics") that avoids the [pitfalls](https://www.nature.com/articles/nmeth.3288 "The fickle P value generates irreproducible results, Halsey et al 2015") of significance testing. It uses familiar statistical concepts: means, mean differences, and error bars. More importantly, it focuses on the effect size of one's experiment/intervention, as opposed to a false dichotomy engendered by _P_ values.

An estimation plot has two key features.

1. It **presents all datapoints** as a swarmplot, which orders each point to display the underlying distribution.

2. It presents the **effect size** as a **bootstrap 95% confidence interval** on a **separate but aligned axes**.

<insert showpiece here>

## Installation

Currently, this package is only available on Github. A CRAN version is forthcoming.

```r
# install.packages("dabestr") ## Does not work; awaiting CRAN acceptance.

devtools::install_github("ACCLAB/dabestr")
```

## Usage

```{r usage}
library(dabestr)

# Performing unpaired (two independent groups) analysis.
unpaired_mean_diff <- dabest(iris, Species, Petal.Width,
                             idx = c("setosa", "versicolor", "virginica"),
                             paired = FALSE)

# Display the results in a user-friendly format.
unpaired_mean_diff

# Produce a Cumming estimation plot.
plot(unpaired_mean_diff)
```

## How to Cite

[**Moving beyond P values: Everyday data analysis with estimation plots**](https://doi.org/10.1101/377978 "Our BioRxiv preprint")

Joses Ho, Tayfun Tumkaya, Sameer Aryal, Hyungwon Choi, Adam Claridge-Chang

## dabest In Other Languages

dabestr is also available in [Python](https://github.com/ACCLAB/DABEST-python "DABEST-Python on Github") and [Matlab](https://github.com/ACCLAB/DABEST-Matlab "DABEST-Matlab on Github").

## Bugs

Please open a [new issue](https://github.com/ACCLAB/dabestr/issues/new). Include a reproducible example (aka [reprex](https://www.tidyverse.org/help/)) so anyone can copy-paste your code and move quickly towards helping you out!

## Contributing

All contributions are welcome. Please fork this Github repo and open a pull request.
