---
title: "Tutorial: Delta-Delta"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Delta-Delta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(dabestr)
```

This vignette documents how `dabestr` is able to compute the calculation of delta-delta, an experimental function that allows the comparison between two bootstrapped effect sizes computed from two independent categorical variables.

Many experimental designs investigate the effects of two interacting independent variables on a dependent variable. The delta-delta effect size lets us distill the net effect of the two variables. To illustrate this, let’s delve into the following problem.

Consider an experiment where we test the efficacy of a drug named `Drug` on a disease-causing mutation `M` based on disease metric `Y`. The greater value `Y` has the more severe the disease phenotype is. Phenotype `Y` has been shown to be caused by a gain of function mutation `M`, so we expect a difference between wild type (`W`) subjects and mutant subjects (`M`). Now, we want to know whether this effect is ameliorated by the administration of `Drug` treatment. We also administer a placebo as a control. In theory, we only expect `Drug` to have an effect on the `M` group, although in practice many drugs have non-specific effects on healthy populations too.

Effectively, we have 4 groups of subjects for comparison.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
df <- data.frame(
  `s` = c("Drug", "Placebo"),
  `Wild type` = c("$X_D, W$", "$X_P, W$"),
  `Mutant` = c("$X_D, M$", "$X_P, M$")
)
colnames(df) <- c(" ", "Wild type", "Mutant")
knitr::kable(df, escape = FALSE) %>%
  kableExtra::column_spec(1, bold = TRUE) %>%
  kableExtra::column_spec(1:2, border_right = TRUE)
```

There are 2 `Treatment` conditions, `Placebo` (control group) and `Drug` (test group). There are 2 `Genotypes`: `W` (wild type population) and `M` (mutant population). In addition, each experiment was done twice (`Rep1` and `Rep2`). We shall do a few analyses to visualise these differences in a simulated dataset.

```{r setup, eval = FALSE}
library(dabestr)
```

## Create dataset for demo
```{r}
set.seed(12345) # Fix the seed so the results are replicable.
# pop_size = 10000 # Size of each population.
N <- 20 # The number of samples taken from each population

# Create samples
placebo <- rnorm(N / 2, mean = 4, sd = 0.4)
placebo <- c(placebo, rnorm(N / 2, mean = 2.8, sd = 0.4))
drug <- rnorm(N / 2, mean = 3, sd = 0.4)
drug <- c(drug, rnorm(N / 2, mean = 2.5, sd = 0.4))

# Add a `Genotype` column as the second variable
genotype <- c(rep("M", N / 2), rep("W", N / 2))

# Add an `id` column for paired data plotting.
id <- 1:N

# Add a `Rep` column as the first variable for the 2 replicates of experiments done
Rep <- rep(c("Rep1", "Rep2"), N / 2)

# Combine all columns into a DataFrame.
df <- tibble::tibble(
  Placebo = placebo,
  Drug = drug,
  Genotype = genotype,
  ID = id,
  Rep = Rep
)

df <- df %>%
  tidyr::gather(key = Treatment, value = Measurement, -ID, -Genotype, -Rep)
```

```{r}
knitr::kable(head(df))
```

## Loading Data
To make a delta-delta plot, you need to simply set `delta2 = TRUE` in the `load()` function. `colour` will be used to determine the colour of dots for scattered plots or the colour of lines for slopegraphs. The `experiment` input will be used to specify grouping of the data.

For delta-delta plots, the `idx` is a non-compulsory input.

## Unpaired Data

```{r, eval = FALSE}
unpaired_delta2 <- load(df,
  x = Genotype, y = Measurement,
  experiment = Treatment, colour = Genotype,
  delta2 = TRUE
)
```

```{r, echo = TRUE}
unpaired_delta2 <- load(df,
  x = Genotype, y = Measurement,
  experiment = Treatment, colour = Genotype,
  delta2 = TRUE,
  experiment_label = c("Placebo", "Drug"),
  x1_level = c("W", "M")
)
```

The above function creates the following `dabest` object:

```{r}
print(unpaired_delta2)
```

We can quickly check out the effect sizes:

```{r}
unpaired_delta2.mean_diff <- mean_diff(unpaired_delta2)

print(unpaired_delta2.mean_diff)
```

```{r}
dabest_plot(unpaired_delta2.mean_diff)
```

In the above plot, the horizontal axis represents the `Genotype` condition and the dot colour is also specified by `Genotype`. The left pair of scattered plots is based on the `Placebo` group while the right pair is based on the `Drug` group. The bottom left axis contains the two primary deltas: the `Placebo` delta and the `Drug` delta. We can easily see that when only the placebo was administered, the mutant phenotype is around 1.23 [95%CI 0.948, 1.52]. This difference was shrunken to around 0.326 [95%CI 0.0934, 0.584] when the drug was administered. This gives us some indication that the drug is effective in amiliorating the disease phenotype. Since the `Drug` did not completely eliminate the mutant phenotype, we have to calculate how much net effect the drug had. This is where delta-delta comes in. We use the `Placebo` delta as a reference for how much the mutant phenotype is supposed to be, and we subtract the `Drug` delta from it. The bootstrapped mean differences (delta-delta) between the `Placebo` and `Drug` group are plotted at the right bottom with a separate y-axis from other bootstrap plots. This effect size, at about -0.903 [95%CI -1.28, -0.513], is the net effect size of the drug treatment. That is to say that treatment with drug A reduced disease phenotype by 0.903.

Mean difference between mutants and wild types given the placebo treatment is:

$$\Delta_1 = \bar{X}_{P,M}-\bar{X}_{P,W}$$

Mean difference between mutants and wild types given the drug treatment is:

$$\Delta_2 = \bar{X}_{D,M}-\bar{X}_{D,W}$$
The net effect of the drug on mutants is:

$$\Delta_\Delta = \Delta_1 - \Delta_2$$
where $\bar{X}$ is the sample mean, $\Delta$ is the mean difference.

## Specifying Grouping for Comparisons
In the example above, we used the convention of "test - control' but you can manipulate the orders of experiment groups as well as the horizontal axis variable by setting `experiment_label` and `x1_level`.

```{r}
unpaired_delta2_specified.mean_diff <- load(df,
  x = Genotype, y = Measurement,
  experiment = Treatment, colour = Genotype,
  delta2 = TRUE,
  experiment_label = c("Drug", "Placebo"),
  x1_level = c("M", "W")
) %>%
  mean_diff()

dabest_plot(unpaired_delta2_specified.mean_diff)
```

## Paired Data
The delta - delta function also supports paired data, which is useful for us to visualise the data in an alternate way. Assuming that the placebo and drug treatment were done on the same subjects, our data is paired between the treatment conditions. We can specify this by using `Treatment` as `x` and `Genotype` as `experiment`, and we further specify that `id_col` is `ID`, linking data from the same subject with each other. Since we have done two replicates of the experiments, we can also colour the slope lines according to `Rep`.

Although the `idx` is a non-compulsory parameter, it is still possible to have it as an input to adjust the order as opposed to using `experiment_label` and `x1_level`.

```{r}
paired_delta2.mean_diff <- load(df,
  x = Treatment, y = Measurement,
  experiment = Genotype, colour = Rep,
  delta2 = TRUE,
  idx = list(
    c("Placebo W", "Drug W"),
    c("Placebo M", "Drug M")
  ),
  paired = "baseline", id_col = ID
) %>%
  mean_diff()

dabest_plot(paired_delta2.mean_diff,
  raw_marker_size = 0.5, raw_marker_alpha = 0.3
)
```

We see that the drug had a non-specific effect of `r format(paired_delta2.mean_diff[["boot_result"]][["difference"]][[1]], digits=3)` [95%CI `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_low"]][[1]], digits=3)` , `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_high"]][[1]], digits=3)`] on wild type subjects even when they were not sick, and it had a bigger effect of `r format(paired_delta2.mean_diff[["boot_result"]][["difference"]][[2]], digits=3)` [95%CI `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_low"]][[2]], digits=3)` , `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_high"]][[2]], digits=3)`] in mutant subjects. In this visualisation, we can see the delta-delta value of `r format(paired_delta2.mean_diff[["boot_result"]][["difference"]][[3]], digits=3)` [95%CI `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_low"]][[3]], digits=3)` , `r format(paired_delta2.mean_diff[["boot_result"]][["bca_ci_high"]][[3]], digits=3)`] as the net effect of the drug accounting for non-specific actions in healthy individuals

Mean difference between drug and placebo treatments in wild type subjects is:

$$\Delta_1 = \bar{X}_{D,M}-\bar{X}_{P,W}$$

Mean difference between drug and placebo treatments in mutant subjects is:

$$\Delta_2 = \bar{X}_{D,M}-\bar{X}_{P,W}$$
The net effect of the drug on mutants is:

$$\Delta_\Delta = \Delta_2 - \Delta_1$$
where $\bar{X}$ is the sample mean, $\Delta$ is the mean difference.

## Connection to ANOVA
The configuration of comparison we performed above is reminiscent of a two-way ANOVA. In fact, the delta - delta is an effect size estimated for the interaction term between `Treatment` and `Genotype`. Main effects of `Treatment` and `Genotype`, on the other hand, can be determined by simpler, univariate contrast plots.

## Omitting Delta-delta Plot
If for some reason you don't want to display the delta-delta plot, you can easily do so by

```{r}
dabest_plot(unpaired_delta2.mean_diff, show_delta2 = FALSE)
```

## Other Effect Sizes
Since the delta-delta function is only applicable to mean differences, plots of other effect sizes will not include a delta-delta bootstrap plot.

```{r}
# cohens_d(unpaired_delta2)
```

## Statistics
You can find all outputs of the delta - delta calculation by assessing the column named `boot_result` of the `dabest_effectsize_obj`.

```{r}
print(unpaired_delta2.mean_diff$boot_result)
```

If you want to extract the permutations, permutation test’s p values, the statistical tests and the p value results, you can access it with the columns `permutation_test_results`, `pval_permtest`, `pval_for_tests` and `pvalues` respectively.

P values for permutation tests `pval_permtest` (and the permutation calculations and results accessed by running the commented out session).
```{r}
# print(unpaired_delta2.mean_diff$permtest_pvals$permutation_test_results)
print(unpaired_delta2.mean_diff$permtest_pvals$pval_permtest)
```

An representative p value for statistical tests (`pval_for_tests`)

```{r}
print(unpaired_delta2.mean_diff$permtest_pvals$pval_for_tests)
```

Statistical test results and `pvalues`.
```{r}
print(unpaired_delta2.mean_diff$permtest_pvals$pvalues)
```
