---
title: "Tutorial: Basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette documents the basic functionalities of dabestr. It illustrates the 
order in which the functions are meant to be used - procedurally. 

The dataset is first processed into the dabestr format via `load()`. The effect sizes are then 
calculated via `effect_size()`. Lastly, the estimation plots are produced via
`dabest_plot()`

```{r setup, warning = FALSE, message = FALSE}
library(dabestr)
```

## Create dataset for demo
Here, we create a dataset to illustrate how dabest functions. In this dataset, 
each column corresponds to a group of observations.
```{r}
set.seed(12345) # Fix the seed so the results are replicable.
# pop_size = 10000 # Size of each population.
N <- 20

# Create samples
c1 <- rnorm(N, mean = 3, sd = 0.4)
c2 <- rnorm(N, mean = 3.5, sd = 0.75)
c3 <- rnorm(N, mean = 3.25, sd = 0.4)

t1 <- rnorm(N, mean = 3.5, sd = 0.5)
t2 <- rnorm(N, mean = 2.5, sd = 0.6)
t3 <- rnorm(N, mean = 3, sd = 0.75)
t4 <- rnorm(N, mean = 3.5, sd = 0.75)
t5 <- rnorm(N, mean = 3.25, sd = 0.4)
t6 <- rnorm(N, mean = 3.25, sd = 0.4)

# Add a `gender` column for coloring the data.
gender <- c(rep("Male", N / 2), rep("Female", N / 2))

# Add an `id` column for paired data plotting.
id <- 1:N

# Combine samples and gender into a DataFrame.
df <- tibble::tibble(
  `Control 1` = c1, `Control 2` = c2, `Control 3` = c3,
  `Test 1` = t1, `Test 2` = t2, `Test 3` = t3, `Test 4` = t4, `Test 5` = t5, `Test 6` = t6,
  Gender = gender, ID = id
)

df <- df %>%
  tidyr::gather(key = Group, value = Measurement, -ID, -Gender)
```

Note that we have 9 groups (3 Control samples and 6 Test samples). Our dataset also has a non-numerical column indicating gender, and another column indicating the identity of each observation.

This is known as a ‘long’ dataset. See this [writeup](https://simonejdemyr.com/r-tutorials/basics/wide-and-long/) for more details.
```{r}
head(df)
```

## Loading Data (Step 1)
Before we create estimation plots and obtain confidence intervals for our effect sizes, we need to load the data and the relevant groups.

We simply supply the DataFrame to `load()` along with x and y - the columns in the DataFrame that contains the treatment groups and measurement values respectively. We also must supply the two groups you want to compare in the `idx` argument as a vector or list.
```{r}
two_groups_unpaired <- load(df,
  x = Group, y = Measurement,
  idx = c("Control 1", "Test 1")
)
```

Printing this `dabestr` object gives you a gentle greeting, as well as the comparisons that can be computed.
```{r}
print(two_groups_unpaired)
```

### Changing statistical parameters
You can change the width of the confidence interval that will be produced by manipulating the `ci` argument.
```{r}
two_groups_unpaired_ci90 <- load(df,
  x = Group, y = Measurement,
  idx = c("Control 1", "Test 1"), ci = 90
)
```

```{r}
print(two_groups_unpaired_ci90)
```

## Effect sizes (Step 2)
`dabestr` now features a range of effect sizes:

- the mean difference (`mean_diff()`)
- the median difference (`median_diff()`)
- Cohen’s d (`cohens_d()`)
- Hedges’ g (`hedges_g()`)
- Cliff’s delta (`cliffs_delta()`)

The output of the `load()` function, a `dabest` object, is then passed into 
these `effect_size()` functions as a parameter.

```{r}
two_groups_unpaired.mean_diff <- mean_diff(two_groups_unpaired)

print(two_groups_unpaired.mean_diff)
```

For each comparison, the type of effect size is reported (here, it’s the “unpaired mean difference”). The confidence interval is reported as: _[confidenceIntervalWidth LowerBound, UpperBound]_

This confidence interval is generated through bootstrap resampling. See Bootstrap Confidence Intervals for more details.

### P-values and statistical tests
Permutation P values are only provided to allow analysts to satisfy a customary requirement of scientific journals. DABEST's provision of P values does not constitute an endorsement of P values or null-hypothesis significance testing (NHST). If users need to include these in a study, we recommend that they (1) avoid performing NHST, i.e. do not compare P to an alpha, (2) never refer to the P values in the Results text, and (3) state in their Methods section that "No null-hypothesis significance testing was performed; P values are provided for legacy purposes only."

## Producing estimation plots (Step 3)
To produce a **Gardner-Altman estimation plot**, simply use the `dabest_plot()`. You can read more about its genesis and design inspiration at Robust and Beautiful Statistical Visualization. 

`dabest_plot()` only requires one compulsory parameter to run: the `dabest_effectsize_obj` obtained from the `effect_size()` function. This means you can quickly create plots for different effect sizes easily.
```{r}
dabest_plot(two_groups_unpaired.mean_diff)
# dabest_plot(two_groups_unpaired.hedges_g)
```

Instead of a Gardner-Altman plot, you can produce a **Cumming estimation plot** by setting `float_contrast = FALSE` in the `dabest_plot()` function This will plot the bootstrap effect sizes below the raw data, and also displays the the mean (gap) and ± standard deviation of each group (vertical ends) as gapped lines. This design was inspired by Edward Tufte’s dictum to maximise the data-ink ratio.

```{r, eval = FALSE}
dabest_plot(two_groups_unpaired.mean_diff,
  float_contrast = FALSE,
  contrast_ylim = c(-0.3, 1.3)
)
```

```{r, echo = FALSE}
pp_plot <- dabest_plot(two_groups_unpaired.mean_diff,
  float_contrast = FALSE,
  contrast_ylim = c(-0.3, 1.3)
)

cowplot::plot_grid(
  plotlist = list(NULL, pp_plot, NULL),
  nrow = 1,
  ncol = 3,
  rel_widths = c(2.5, 5, 2.5)
)
```

The `dabestr` package also implements a range of estimation plot designs aimed at depicting common experimental designs.

The **multi-two-group estimation plot** tiles two or more Cumming plots horizontally, and is created by passing a _nested list_ to `idx` when `load()` is first invoked.

Thus, the lower axes in the Cumming plot is effectively a [forest plot](https://en.wikipedia.org/wiki/Forest_plot), used in meta-analyses to aggregate and compare data from different experiments.

```{r}
multi_2group <- load(df,
  x = Group, y = Measurement,
  idx = list(
    c("Control 1", "Test 1"),
    c("Control 2", "Test 2")
  )
)
multi_2group %>%
  mean_diff() %>%
  dabest_plot()
```

The **shared control plot** displays another common experimental paradigm, where several test samples are compared against a common reference sample.

This type of Cumming plot is automatically generated if the vector passed to `idx` has more than two data columns.
```{r}
shared_control <- load(df,
  x = Group, y = Measurement,
  idx = c(
    "Control 1", "Test 1", "Test 2", "Test 3",
    "Test 4", "Test 5", "Test 6"
  )
)

print(shared_control)
```

```{r}
shared_control.mean_diff <- mean_diff(shared_control)

print(shared_control.mean_diff)
```

```{r}
dabest_plot(shared_control.mean_diff)
```

`dabestr` thus empowers you to robustly perform and elegantly present complex visualizations and statistics.
```{r}
multi_groups <- load(df,
  x = Group, y = Measurement,
  idx = list(
    c("Control 1", "Test 1"),
    c("Control 2", "Test 2", "Test 3"),
    c("Control 3", "Test 4", "Test 5", "Test 6")
  )
)

print(multi_groups)
```

```{r}
multi_groups.mean_diff <- mean_diff(multi_groups)

print(multi_groups.mean_diff)
```

```{r}
dabest_plot(multi_groups.mean_diff)
```
 
## Using wide datasets
`dabestr` does not currently support the use of 'wide' data. To convert datasets from 'wide' to 'long'/'tidy', consider taking a look at [gather()](https://tidyr.tidyverse.org/reference/gather.html) as part of the [tidyr](https://tidyr.tidyverse.org) package.
